---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `ggstats`: extension to `ggplot2` for plotting stats

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/larmarange/ggstats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/larmarange/ggstats/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/larmarange/ggstats/branch/main/graph/badge.svg)](https://app.codecov.io/gh/larmarange/ggstats?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/ggstats)](https://CRAN.R-project.org/package=ggstats)
<!-- [![DOI](https://zenodo.org/badge/286680847.svg)](https://zenodo.org/badge/latestdoi/286680847) -->
<!-- badges: end -->

The `ggstats` package provides  suite of functions to plot regression model coefficients ("forest plots") using `ggplot2`..

The suite also includes new statistics to compute proportions, weighted mean and cross-tabulation statistics, as well as new geometries to add alternative background color to a plot.

The original versions of several functions were originally developed within the `{GGally}` package.

## Installation

<!-- To install stable version:

```{r eval=FALSE}
install.packages("ggstats")
```

-->
To install development version:

```{r eval=FALSE}
remotes::install_github("larmarange/ggstats")
```

## A basic example


```{r}
library(ggstats)

data(tips, package = "reshape")
mod_simple <- lm(tip ~ day + time + total_bill, data = tips)
ggcoef_model(mod_simple)
```

## Comparing several models

```{r}
mod1 <- lm(Fertility ~ ., data = swiss)
  mod2 <- step(mod1, trace = 0)
  mod3 <- lm(Fertility ~ Agriculture + Education * Catholic, data = swiss)
  models <- list("Full model" = mod1, "Simplified model" = mod2, "With interaction" = mod3)

  ggcoef_compare(models)
  ggcoef_compare(models, type = "faceted")
```
