---
title: "Geometries for diverging bar plots"
author: Joseph Larmarange
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geometries for diverging bar plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggstats)
library(dplyr)
library(ggplot2)
library(patchwork)
```

*Note :* if you are looking for an all-in-one function to display Likert-type items, please refer to `gglikert()` and `vignette("gglikert")`.

## New positions

Diverging bar plots could be achieved using `position_diverging()` or `position_likert()`.

`position_diverging()` stacks bars on top of each other and centers them around zero (the same number of categories are displayed on each side).

```{r}
base <-
  ggplot(diamonds) +
  aes(y = clarity, fill = cut) +
  theme(legend.position = "none")

p_stack <-
  base +
  geom_bar(position = "stack") +
  ggtitle("position_stack()")

p_diverging <-
  base +
  geom_bar(position = "diverging") +
  ggtitle("position_diverging()")

p_stack + p_diverging
```

`position_likert()` is similar but uses proportions instead of counts.

```{r}
p_fill <-
  base +
  geom_bar(position = "fill") +
  ggtitle("position_fill()")

p_likert <-
  base +
  geom_bar(position = "likert") +
  ggtitle("position_likert()")

p_fill + p_likert
```

By default, the same number of categories is displayed on each side, i.e. if you have 4 categories, 2 will be displayed negatively and 2 positively. If you have an odd number of categories, half of the central category will be displayed negatively and half positively.

The <q>center</q> could be changed with the `cutoff` argument, representing the number of categories to be displayed negatively: `2` to display negatively the two first categories, `2.5` to display negatively the two first categories and half of the third, `2.2` to display negatively the two first categories and a fifth of the third.

```{r}
p_1 <-
  base +
  geom_bar(position = position_diverging(cutoff = 1)) +
  ggtitle("cutoff = 1")

p_2 <-
  base +
  geom_bar(position = position_diverging(cutoff = 2)) +
  ggtitle("cutoff = 2")

p_null <-
  base +
  geom_bar(position = position_diverging(cutoff = NULL)) +
  ggtitle("cutoff = NULL")

p_3.75 <-
  base +
  geom_bar(position = position_diverging(cutoff = 3.75)) +
  ggtitle("cutoff = 3.75")

p_5 <-
  base +
  geom_bar(position = position_diverging(cutoff = 5)) +
  ggtitle("cutoff = 5")

wrap_plots(p_1, p_2, p_null, p_3.75, p_5)
```

## New scales

For a diverging bar plot, it is recommended to use a diverging palette, as provided in the Brewer palettes. Sometimes, the number of available colors is insufficient in the palette. In that case, you could use `pal_extender()` or `scale_fill_extended()`. However, if you use a custom `cutoff`, it is also important to change the center of the palette as well.

Therefore, for diverging bar plots, we recommend to use `scale_fill_likert()`.

```{r}
wrap_plots(
  p_1 + scale_fill_likert(cutoff = 1),
  p_null + scale_fill_likert(),
  p_3.75 + scale_fill_likert(cutoff = 3.75)
)
```

## Improving axes

You may also want have centered axes. That could be easily achieved with `symmetric_limits()`.

You could also use `label_number_abs()` or `label_percent_abs()` to display absolute numbers.

```{r}
wrap_plots(
  p_3.75,
  p_3.75 +
    scale_x_continuous(
      limits = symmetric_limits,
      labels = label_number_abs()
    )
)
```

## New geometries

To facilitate the creation of diverging bar plots, you could use variants of `geom_bar()` and `geom_text()`.

### geom_diverging() & geom_diverging_text()

Let's consider the following plot:

```{r}
ggplot(diamonds) +
  aes(y = clarity, fill = cut) +
  geom_bar(position = "diverging") +
  geom_text(
    aes(
      label =
        label_number_abs(hide_below = 800)
        (after_stat(count))
    ),
    stat = "count",
    position = position_diverging(.5)
  ) +
  scale_fill_likert() +
  scale_x_continuous(
    labels = label_number_abs(),
    limits = symmetric_limits
  )
```

The same could be achieved quicker with `geom_diverging()` and `geom_diverging_text()`.

```{r}
ggplot(diamonds) +
  aes(y = clarity, fill = cut) +
  geom_diverging() +
  geom_diverging_text(hide_below = 800) +
  scale_fill_likert() +
  scale_x_continuous(
    labels = label_number_abs(),
    limits = symmetric_limits
  )
```

### geom_likert() & geom_likert_text()

`geom_likert()` and `geom_likert_text()` works similarly. `geom_likert_text()` takes advantages of `stat_prop()` for computing the proportions to be displayed (see `vignette("stat_prop")`).

```{r}
ggplot(diamonds) +
  aes(y = clarity, fill = cut) +
  geom_likert() +
  geom_likert_text(hide_below = 0.10) +
  scale_fill_likert() +
  scale_x_continuous(
    labels = label_percent_abs()
  )
```

### geom_pyramid() & geom_pyramid_text()

Finally, `geom_pyramid()` and `geom_pyramid_text()` are variations adapted to display an age-sex pyramid. It uses proportions of the total.

```{r}
d <- Titanic |> as.data.frame()
ggplot(d) +
  aes(y = Class, fill = Sex, weight = Freq) +
  geom_pyramid() +
  geom_pyramid_text() +
  scale_x_continuous(
    labels = label_percent_abs(),
    limits = symmetric_limits
  )
```
